Fast and Flexible Binary Instrumentation for the Kernel

### 2013-12-13 VEE-14/paper.tex
Updated section 4.

\maketitle
\subsection*{Abstract}
%Kernel modules extend the functionality of operating systems. They represent the bulk of new kernel code in development and contain the majority of new kernel bugs. Unfortunately, analyzing and debugging modules is very challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which makes static analysis intractable. Existing dynamic analysis tools capable of instrumenting modules are not suited toward module analysis. They either impose unnecessary performance overheads, lack flexibility and are too coarse-grained, or do not provide a means of understanding the ``big picture'' view of what an arbitrary module does as it executes.

%We created Granary to address the challenges of module analysis. Granary is a framework for creating flexible and efficient tools that analyze or debug arbitrary, binary Linux kernel modules. Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. Granary makes it easy to create flexible tools by supporting context-aware runtime code specialization, and by integrating high-level static analysis information with low-level instruction manipulation. Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.

@@

%Operating system (OS) kernels present an apparently insane environment for a transparent binary instrumentation framework.  For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state~\cite{DRK}. Preserving interrupt transparency, however, requires that interrupts be delayed until the end of any injected instrumentation instructions.  Together, these requirements make it very difficult to write instrumentation tools.  The insane environment of OS kernels also creates several opportunities for analysis tools. For example, interrupts can be ``tamed'' and put to new uses; privileged hardware features can be used; and the kernel ABI and source code can be relied upon to be a good predictor of module behavior at the kernel/module boundary. We also found that the kernel is largely insensitive to transparency issues, allowing us to relax transparency in exchange for both improved performance and greater visibility into the execution of instrumented code.  This increased visibility is particularly important for analysis and debugging tools.

%The challenges and opportunities of the kernel environment motivated the creation of Granary. Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analyzing kernel modules. Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}. Granary allows developers to create flexible and dynamic analysis tools for arbitrary Linux kernel module binaries. Granary has two main goals: \begin{inparaenum}[i)]
%	\item make it easy to create efficient, low-level analysis tools; and
%	\item provide high-level static analysis information to the analysis tools.
%\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts. Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code. By meeting these goals, Granary tools are able to do things like  figure out which data structure fields are associated with each memory access.

%Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
%	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU};
%	\item they do not allow for flexible runtime code specialization; and
%	\item they do not provide static type information to the instrumented code.
%\end{inparaenum}

@@

%This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.

%This is challenging because Granary must be comprehensive: all code that is meant to be instrumented should be instrumented. In other words, relinquishing control to execute kernel code natively must not cause Granary to miss the execution of any module code.

@@

%Also, the runtime checks needed to multiplex between different versions of instrumented code are orthogonal to the true goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.

%Using an existing system, 
% get around the problem of needing different instrumentation in different contexts, existing systems need to create as many versions of the code as there are contexts. Runtime checks must also be inserted into the code that coordinate with the runtime to track and detect the current execution context and jump to the correct version of the code. Runtime checks and state tracking impose unnecessary complications to tool-authors. The added code needed to query the current context and switch to a different version of the instrumentation is code that is not actively achieving the primary goal of the tool, and is instead making up for a defeciency in the runtime system.  
%are orthogonal to the main goals of an analysis tool:
% Aside from the additional ineffeciency of these runtime checks, there is also the 
% However, generating many differently instrumented versions of the code introduces bloat, especially when not all versions are well exercised. Also, generating many different versions of the code introduces state tracking and switching costs.
% This imposes unnecessary performance overheads because of the need to additional instrumentation that is unrelated to 
% due to having to perform runtime detection and swichting
% to the right version of the code for every basic block or every instruction.
% according to \emph{all} contexts. Instrumenting code according to all contexts imposes unnecessary performance overheads and requires state tracking.
%,  dynamically tracking the current context, and dynamically choosing which 
% commit to instrumenting code to handle \emph{all} contexts. For example, if

@@

%There are similar checks for write-side pointers in write-side critical sections.
% The instrumentation performed inside of read- and write-side critical sections is heavy weight, while code running in the third execution context can either run with lightweight or no instrumentation.
%RCUdbg distinguishes between RCU-protected pointers and plain old pointers. There are two kinds of RCU-protected pointer (read-only, and read/write), and every access to an RCU-protected pointer is checked. The checks performed on an access to an RCU-protected pointer are context-specific. That is, a read-only pointer can only be read from within an RCU read-side critical section. Read/write pointers have similar restrictions. RCUdbg needs to know in which context some code is executing, i.e. in a read-side or write-side critical section, or outside of all critical section, and change instrumentation to check for different kinds of bugs based on the context.
%A read/write pointer has different restrictions. In read-side critical sections, a read/write pointer can only be accessed by means of a special API function. In write-side critical sections, a read/write pointer can only be written to 
% One kind of RCU-protected pointer can be read from within ``RCU read-side critical sections", while another kind can (almost) arbitrarily accessed, but only inside of an ``RCU write-side critical section". Outside of these two kinds of sections, accessing RCU-protected pointers is illegal. The distinction between different contexts of code (read-side, write-side, outside), and different kins of pointers makes
%Creating a tool to catch these kinds of bugs with an existing DBT system would be difficult because that tool would need to  
%For example, Granary's Read-Copy-Update (RCU) debugging tool, \toolname{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API. Correct usage of the RCU API allows programmers to create scalable, shared data structures that permit concurrent access and modification. The ``trick" of RCU is that readers might not see the latest write to the data structure, and writers must synchronize to modify the data structure. \toolname{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently. In read-side critical sections, \toolname{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data. In write-side critical sections, \toolname{RCUdbg} detects when writers write to shared data without using the correct API. 
%In reader-side code
%For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs). The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency. Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function. This is challenging with existing DBT systems because some optimized functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs. Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.

@@

%Granary records 
%Policy information is recorded inside of Granary's

% instruction by telling Granary what policy to use to instrument the code 
%, and when the \texttt{call} is eventually executed, 
%This simplifies tool creation becase 
%By propagating policy information through control-flow transferring instructions, 
%Therefore, we can know at translation and instrumentation time that
% No additional state needs to be maintained by \texttt{RCUdbg} to determine which version of the basic block to execute.
%Any control-flow instructions the transfer control to other basic blocks that aren't the \texttt{rcu\_read\_unlock} function must also 
%That is, the failure of existing systems is that they have no built-in support for maintaining multiple versions of code. Tools built on existing systems must emulate versioning by duplicating their instrumentation for each desired version, and add code that executes the right version at the right time.  
%, as would be done using existing systems, 
%What makes Granary's runtime code specialization efficient is that it ties into Granary's \texttt{just-in-time} instrumentation model. That is, Granary only instruments code that will execute. Code that never executes is neither translated nor instrumented. Also, only one version of a basic block is generated at a time. Therefore, even though \toolname{RCUdbg} defines three policies, only
%If some native code is about to execute, then Granary will only instrument it with the policy associated with 
%, and an instrumentation policy is only applied to code if that code will execute within the context associated with that context.
%defines three different instrumentation policies--one for each execution context--and declares switches 
% instrumentation policies to 
%only needs to define three instrumentation policies (one for each execution context), and declare 
%different versions of the code are generated \emph{just-in-time}. Therefore, if the \texttt{memset} function is never executed inside of a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context. Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated. 
%However, unlike other systems, no runtime check that chooses to dispatch to one version of \texttt{memset} or another depending on the runtime context. Instead, 
% if some code never executes in a particular context, then Granary will not generate an instrumented version of that code for that context.  The second benefit of this approach is that it is stateless, and requires no runtime checks to switch between different versions of the code depending on the execution context. 
%Granary tracks the current execution context with zero runtime overhead by 
%In a Granary tool, each tracked execution context is associated with its own instrumentation policy. An instrumentation policy decides how to instrument code that will execute within the associated context. 
%Different versions of the code are instrumented \emph{just-in-time}. That is, there will 
%Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks. This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, reentrancy, and the inability to predict later execution contexts at translation time).
% In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks.  In the case of our CFG tool, two policies track the two execution contexts: $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$. $P_{\mathit{call\_entry}}$ applies to the first execution of the first basic block of each function, and connects functions in the inter-procedural CFG.  $P_{\mathit{after\_entry}}$ applies to all other executed basic blocks within each function, including re-executions of the first basic block. $P_{\mathit{after\_entry}}$ connects basic blocks in the intra-procedural CFG. 
%Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information. Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviors that are only available to low-level runtime instrumentation tools. Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time. The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}. Granary statically analyzes the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact. Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}. Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.
%For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules. This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file. From these models, we can construct behavioral patterns about how modules use kernel data structures and functions. These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behavior \cite{LXFI}.

@@

% \Cref{sec:modes} describes how to efficiently implement mixed-mode execution. \Cref{sec:policies} describes a new method for dynamically changing how code is instrumented. \Cref{sec:reify} describes how to integrate static analysis information into a DBT system. Finally, \Cref{sec:eval} evaluates Granary's performance.

@@

%Granary translates and instruments code at one of three default granularities: kernel, module, and function (also called probe-based instrumentation). 
%
%At runtime, the granularity of instrumentation can be altered based on the execution context. For example, a tool analyzing modules for bugs can choose not to instrument a particular function (e.g. whitelist the function) if that function is known to not contain bugs, or known to generate many false-positives.
%
%\paragraph{Kernel}
%Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively. 
%
%\paragraph{Module}
%
%
%\paragraph{Function}
%
%Granary can interpose on the execution of specific kernel functions in three ways:
%\begin{enumerate}
%	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code. A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code). We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
%	\item {\bf Substituting:} Granary can substitutes the execution of a kernel function, even when that function is executed by native kernel code. This is useful when tools only instrument part of the kernel (e.g. modules or functions) but require visibility on specific kernel events happening outside of the instrumented code (e.g. particular memory allocations).
%	\item {\bf Instrumented Wrapping/Substituting:} Granary can optionally instrument wrapped or substituted code. This is useful when tools want to observe function arguments/return values using high-level C/C++ code, but where that function must be instrumented for visibility.
%\end{enumerate}

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch (\texttt{jCC}), \texttt{ret}, or a \texttt{jmp} instruction. Unlike DRK and btkernel, Granary does not terminate basic blocks at \texttt{call} instructions. This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}). Granary takes control of execution starting from ``natural" program entrypoints (e.g. interrupt vectors, system call entrypoints, module initializers). Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.

@@

%Granary code executes with interrupts disabled and depends on CPU-private data to avo
%Granary's translation mechanism depends on CPU-private data to avoid contention when two or more processors simultaneously execute Granary code.
%because kernel stacks are typically small (4K to 8K in size), and because context-switches into Granary can occur at arbitrary program points, Granary cannot risk 
%and a context switch into Granary might occur at any point. If most of the stack is already used by current thread before it yields control to Granary then

@@

Associated with each basic block is meta-data that describes the basic block. Basic block meta-data includes: \begin{itemize}
	\item The length in bytes of the original and translated basic blocks, and the address of the first instruction in the original basic block and the instrumented basic block. This information helps provide a rich debugging interface for instrumentation tool authors. Granary is unique among DBT tools in that the debugging experience is actually pleasant. For example, Granary has a built-in debugging mechanism for stepping back in time through previously executed basic blocks, as well as querying past register state.
	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block. This kernel-specific meta-data is used by Granary to implement transparent page fault exception handling (\Cref{sec:interrupt_return_address_transparency}).
	\item Ranges of instructions that must be executed atomically. Granary uses this information to implement selective interrupt delaying (\Cref{sec:interrupt_delay}).
	\item The allocator used to allocate the instructions of the basic block. Granary uses this information to implement its trace-allocation optimization (\Cref{sec:trace_alloc}).
	\item The instrumentation policy used to instrument the basic block. Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines. For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
\end{itemize} %
%Granary uses basic block meta-data for trace allocation (\Cref{sec:trace_alloc}), interrupt delaying (\Cref{sec:interrupt_delay}), and policy-specific interrupt handling.

Basic block meta-data can be also be extended by instrumentation tools. This makes developing instrumentation tools that record per-basic block statistics very easy to make. For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts (\Cref{sec:cfg}). This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization. This meta-data is specified by \toolname{cfg} and automatically allocated by Granary.

%Basic block meta-data can also contain Granary's intermediate representation for x86-64 instructions. This is a unique feature that enables instrumentation tools to implement their own form of hot-code patching (among other things), as the intermediate representation remains tied to an emitted x86-64 instruction once the block is committed to the code cache.

%Unlike any other instrumentation tool, Granary allows the x86-64 instruction intermediate representation

%Another

%novelty of Granary is that basic block meta-data enables a good

%Basic block meta-data is queried and extended by Granary instrumentation tools.

%// TODO!

%Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block. This enables policy-specific interrupt handling. That is, when a basic block instrumented by a specific policy is interrupted, that policy can choose how to handle the interrupt. For example, Granary's address watchpoints framework can instrument code using a lightweight, \toolname{null} instrumentation.

\subsection{Transparency}\label{sec:transparency}
An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, Granary exposes the following two artifacts: code cache addresses as function and interrupt return addresses.

@@

%performing a \texttt{ret} requires an additional \texttt{jmp} to 
%Neither of DRK's nor btkernel's approaches enable native-like return address prediction by the CPU.
%and then jumping to the next basic block. This in turn . Btkernel's approach to procedure calls introduces similar inefficiencies, despite the fact that does not provide return address transparency, because it also emulates procedure calls, but by manually pushing the instrumented return address onto the stack instead of the native one.
%Granary's approach improves on DRK and btkernel by allowing 
%Better return-address prediction by the processor.

@@

	%longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
\end{enumerate}

@@

%\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions. This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behavior under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.

@@

%Another concern is that instrumentation itself is not re-entrant. For example, instrumentation instructions added by a tool that access CPU-private information

@@

%Therefore, if a context switch into Granary occurs because the next basic block must be found/translated, then the address of that next basic block must be stored \emph{somewhere} in order for execution to eventually reach that basic block when the context switches back to the code cache. But context switching is not allowed to perturb the register state of the instrumented program, and so the target address must be resolved by another means.

%Like btkernel, Granary solves this problem of figuring out how to get to where the code needs to go  by generating ``edge'' code for each direct control-flow instruction (e.g. \texttt{call}, \texttt{jmp}, \texttt{jcc}).

%Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code. Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtime call stack. Entry edge is responsible for indirect branch lookup, and will indirectly jump to the exit edge code for a given basic block. Exit edge code is generated for each basic block that is targeted by an indirect CTI. Exit edge code restores the register state saved by the entry edge code, and then directly jumps to the beginning of the targeted basic block.

@@

%If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior. We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers. Because of interrupt delaying, the spilling code was sometimes repeated many times, whereas the restoring was only executed once. This led to hard-to-debug kernel crashes.

@@

% and re-issues the interrupt to Granary. Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.\footnote{Leaving interrupts disabled is required to ensure that executing the copied instructions in the CPU-private buffer is reentrant. Granary performs some sanity checks on the instructions being copied to ensure that they don't re-enable interrupts.} When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame. At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.

%while not supporting delaying has value because not all instrumentation is non-reentrant. By default, Granary does not delay interrupts. However, instrumentation tools can declare \emph{interrupt delay ranges}:

%, but introduces implementation complexities and high overheads for interrupt-heavy workloads. Precise interrupt delivery is one of the ways which DRK maintains transparency: interrupt handlers only 
%The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder. For example, Granary's \toolname{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses. When necessary, \toolname{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code. This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after. DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.

@@

\subsubsection{Direct Control Flow}
% Code block showing an example JZ (jump if ZF=1).
\lstset{language=[x64]Assembler}
\newsavebox\nativejcc
\begin{lrbox}{\nativejcc}
\begin{minipage}[b]{4cm}
\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
cmp $0x0, %rax
jz value_is_zero 
...
( fall through )
\end{lstlisting}
\end{minipage}
\end{lrbox}

@@

% and is responsible for transferring control from a basic block ($B_{pred}$) to Granary so that the next basic block ($B_{succ}$) can be found/translated, and then connected ($B_{pred} \underset{\texttt{jz}}{\to} B_{succ}$) via hot code patching.

@@

%Indirect branch lookup through a global, fixed-size hash table. A basic block making an indirect control flow transfer jumps to an in-edge that knows where to find the indirect CTI target (e.g. in a specific register or memory location). The in-edge then uses the target as an index into the hash table, and jumps to the target address stored in the indexed hash table entry. That entry will either be the address of an out-edge, or of Granary's indirect branch resolver. Each out-edge is specific to a target basic block, and first checks if the intended jump target matches the out-edge's target. If so, then control is transfered to the target basic block, otherwise control is transfered either to Granary or to the next out-edge. Granary chains together out-edges through conditional jumps just as

%Hash table entries are code addresses of out-edge code stub. Each out-edge code stub is specific to a target basic block, and is responsible for checking if the intended control-flow target is the out-edge's target or not. If not, the out-edge either directly jumps to the next out-edge whose target also fits into the same hash table bucket, or jumps to Granary to determine the target basic block.

@@

%If an entry points to Granary, then no as-of-yet resolved indirect jump target hashes to that entry.

%Granary implements separate chaining through code (i.e. control flow jumps, as opposed to date, e.g. linked list pointers)
%In the event that one-or-more indirect jump targets do hash to an existing entry
%In-edge code is responsible for 
% to first search for their targets in a global, fixed-size hash table, and failing that, context-switch into Granary to perform a full lookup and potential translation.
%using a fixed-size hash table, where the native target of the indirect

@@

%using internal performance counters within Granary

%Static branch prediction hints (\texttt{likely}, \texttt{unlikely} macros) are heavily used in the Linux kernel source code. Although static branch prediction hints can be specified on individual conditional branch instructions, recent x86-based architectures ignore these hints \cite{AgnerMicroarchitecture}. Therefore, static branch prediction hints at the source code level mainly affect compiler optimizations, including intraprocedural basic block scheduling. Our tracing strategy is based on

%Tr

%We expect and have experimentally verified that the most common execution path through code is straight-line and does not following conditional branches.

%Due to instruction prefetching, we expect that the likely or

%On recent x86-based architectures, these hints only affect the schedule of basic blocks in the intraprocedural control-flow graph \cite{AgnerMicroarchitecture}.

@@

%Padding basic blocks so as to align specific instructions must not change the execution behavior.

@@

%To reduce the overhead of alignment 
%In some cases, basic blocks can be directly connected during basic block translation time. For example, intra-trace CTIs, or CTIs targeting already translated basic blocks. However, not all basic blocks can be connected this way, therefore some runtime 
%Granary connects some basic blocks at translation time (e.g. intra-trace control-flow instructions). This is common when tracing is enabled, but also occurs when a branch targets an already translated basic block. However, 
%Some basic blocks are directly connected at basic block translation time. One example is branches that target other basic blocks within a trace. Another example is a branch to an already translated basic block.
%Other basic blocks are only connected at runtime. For example, if the function \texttt{memset} has never been executed, then a \texttt{call} to \texttt{memset} will be translated to target edge code (\Cref{fig:direct_edge_code}) that, when executed, will translate \texttt{memset}'s first basic block and patch the \texttt{call} to target the translated basic block.
%For example, the targets of function call instruction (\texttt{call}) are resolved just-in-time (unless the target is already known). For each translated \texttt{call} instruction, Granary generates edge code (\Cref{fig:direct_edge_code}) that will find or translate the first basic block of the function being called, and patch the translated \texttt{call} instruction to point to the function's first basic block.
%For example, if a direct branch in one basic block targets another block within the trace under construction, then that 
%Basic blocks are directly connected at block translation time when branches in traces re-enter a trace, or when branches target previously translated basic blocks. If a branch does not target another basic block within its trace, or does not target a previously translated basic block, then the branch will target edge code that will resolve that branches target at runtime.
% However, when a given basic block is translated, not all of its targets are known (e.g. \texttt{call} and \texttt{jCC} targets). Therefore, edge code (\Cref{fig:direct_edge_code}) is generated that translates/resolves a branch target when that branch is first executed, and patches the translated branch to avoi
% The targets of some of these direct CTIs are only resolved at runtime, by executing edge code. As an optimization, jumps to edge code are dynamically patched to transfer control to their intended targeted basic blocks.
%edge code use hot code patching to update a translated CTI to redirect execution

@@

%, and is associated with a private memory pool from which to allocate memory for basic blocks. The net result is that most of the basic blocks eecuted in service of a single system call are allocated from that system call's memory pool.
%The intuition for this approach is that kernel code is not very ``loopy", and so 
%Granary services memory allocation requests for basic block instructions from memory pools. By default, Granary has as many memory pools as there are CPUs. However, when trace allocation is enabled
%Tracing is also beneficial for 
%However, tracing of this sort requires mechnisms for evicting stale basic blocks, 
%Constructing traces of ``hot'' 
%JIT generally results in better performance .
%By default, Granary allocates the memory needed to hold the instructions of a basic block from CPU-private memory pools. Each memory pool serves basic block allocation requests out of fixed-size memory blocks, ranging from 4K to 24K in size.

@@

%re-direct jumps through the system call table

@@

%Directing control through the shadow system call table avoids the overhead of going through the indirect branch lookup (IBL) table (\Cref{sec:ibl}), and reduces the number of potential conflicts in the IBL table.
% whose entries are code cache basic block addresses translated according to the starting instrumentation policy.
%\subsection{Reproducability}
%Hello

\begin{table*}[ht!]
\scriptsize
\begin{tabularx}{\linewidth}{| l | >{\raggedright}p{0.3\linewidth} | X |}
\hline
\bf Granularity & \bf Description & \bf Example Tool \\
\hline
Kernel & All kernel code is instrumented. & Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters. This data is useful for debugging an performance analysis. Applying the \toolname{cfg} tool to the whole kernel allows us to perform profile-guided optimization of kernel code, including a form of de-virtualization. That is, most indirect control flow instructions (\texttt{call}, \texttt{jmp}) within the kernel target only one destination. Fast checks can be inlined into kernel code that use direct control transfer instructions when possible, and fall-back on indirect control transfer instructions when necessary.\\
\hline
Module & All code from a specific kernel module or driver is instrumented. Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code. All non-module kernel code executes with zero overhead. & Granary's Guardrail tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}. A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs. \\
\hline
Function & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} Linux kernel functions. When invoked by native code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} attaches instrumentation. When invoked by instrumented code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} updates the current nesting depth of the read-side critical section. \\

% Granary's \toolname{Irrigator} tool find leaks on module-owned memory (\Cref{sec:function_wrapper}). \toolname{Irrigator} instruments only module code, but also interposes on specific kernel functions (kernel memory allocators and deallocators) to gain visibility into kernel code that would normally execute natively (using module-only instrumentation).\\
\hline
%Basic Block & Individual basic blocks are instrumented on demand. & Granary's address watchpoints framework can be tuned to instrument only code accessing specific data. A general protection fault occurs when native code dereferences a  ``watched address". Granary can transparently handle this fault by returning execution to an instrumented version of the faulting code (where the instrumentation prevents such faults). Native execution can resume at the end of the basic block. \\
%\hline
\end{tabularx}
\caption{\label{tab:granularities}Different granularities of instrumentation supported by Granary.}
\end{table*}

@@

%The closest that an existing system can get to

>>>>

\maketitle
\subsection*{Abstract}
TODO

@@

@@

@@

@@

@@

@@

@@

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch (\texttt{jCC}), \texttt{ret}, or a \texttt{jmp} instruction. Unlike DRK and btkernel, Granary does not terminate basic blocks at \texttt{call} instructions. This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}). Granary takes control of execution starting from ``natural" program entrypoints (e.g. interrupt vectors, system call entrypoints, module initializers). Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as control flow reaches them for the first time.

@@

@@

Associated with each basic block is meta-data that describes the basic block. Basic block meta-data includes: \begin{itemize}
	\item The length in bytes of the original and translated basic blocks, and the address of the first instruction in the original basic block and the instrumented basic block. This information helps provide a rich debugging interface for instrumentation tool authors. Granary is unique among DBT tools in that the debugging experience is actually pleasant. For example, Granary has a built-in reverse debugging mechanism, which allows developers to inspect previously executed basic blocks and past register state.
	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block. This kernel-specific meta-data is used by Granary to implement transparent page fault exception handling (\Cref{sec:interrupt_return_address_transparency}).
	\item Ranges of instructions that must be executed atomically. Granary uses this information to implement selective interrupt delaying (\Cref{sec:interrupt_delay}).
	\item The allocator used to allocate the instructions of the basic block. Granary uses this information to implement its trace-allocation optimization (\Cref{sec:trace_alloc}).
	\item The instrumentation policy used to instrument the basic block. Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines. For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
\end{itemize}

Basic block meta-data can also be extended by instrumentation tools. This makes developing instrumentation tools that record per-basic block statistics very easy to make. For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts (\Cref{sec:cfg}). This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization. This meta-data is specified by \toolname{cfg} and automatically allocated by Granary.

\subsection{Transparency}\label{sec:transparency}
An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, the only artifact Granary exposes is that interrupt and function return addresses can point into the code cache.

@@

@@

@@

@@

@@

@@

@@

@@

\subsubsection{Direct Control Flow}\label{sec:direct_control_flow}
\lstset{language=[x64]Assembler}
\newsavebox\nativejcc
\begin{lrbox}{\nativejcc}
\begin{minipage}[b]{4cm}
\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
cmp $0x0, %rax
jz value_is_zero 
...
( fall through )
\end{lstlisting}
\end{minipage}
\end{lrbox}

@@

@@

@@

@@

@@

@@

@@

@@

@@

\begin{table*}[ht!]
\scriptsize
\begin{tabularx}{\linewidth}{| l | >{\raggedright}p{0.3\linewidth} | X |}
\hline
\bf Granularity & \bf Description & \bf Example Tool \\
\hline
Kernel & All kernel code is instrumented. & Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters. This data is useful for debugging an performance analysis. Applying the \toolname{cfg} tool to the whole kernel allows us to perform profile-guided optimization of kernel code, including a form of de-virtualization. That is, most indirect control flow instructions (\texttt{call}, \texttt{jmp}) within the kernel target only one destination. Fast checks can be inlined into kernel code that use direct control transfer instructions when possible, and fall-back on indirect control transfer instructions when necessary.\\
\hline
Module & All code from a specific kernel module or driver is instrumented. Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code. All non-module kernel code executes with zero overhead. & Granary's \toolname{Guardrail} tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}. A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs. \\
\hline
Function & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} Linux kernel functions. When invoked by native code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} attaches instrumentation. When invoked by instrumented code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} updates the current nesting depth of the read-side critical section. \\
\hline
\end{tabularx}
\caption{\label{tab:granularities}Different granularities of instrumentation supported by Granary.}
\end{table*}

@@
